---
title       : The Mandelbrot Set
subtitle    : Math is Beautiful
author      : kbelsvik
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## The Problem with Math...

- Far to many people perceive math as being an unappealing subject
- There are a variety of reasons for this perception
    - They find it boring or monotonous
    - Lacks emotional appeal
    - Difficult to understand
- Observe this equation below

$$f(Z_{n+1}) = Z_{n}^2 + C$$

- Nothing exciting here, right? But...

---

## ...doesn't have to be a problem

What if I told you that boring old equation is used to draw something like this?

```{r echo=F, cache=T}
#parameters
    step <- 25
    reps <- 400
    n <- 400
    zoom <- 18.25
    x <- -1.71122185076089
    y <- 3.54780878690511e-05
    w <- 4/2^zoom
    h <- 4/2^zoom
    
    #starting point
    C <- complex(imag=rep(seq(y - h/2, y + h/2, length.out=n), each=n),
                 real=rep(seq(x - w/2, x + w/2, length.out=n), n))
    Z <- rep(0, n*n)
    diverge <- rep(0, n*n)
    
    j <- 0
    upper <- 0
    while(j < reps){
        #perform some house keeping after every 'step' repetitions, for performace improvement
        upper <- ifelse(upper+step < reps, upper+step, reps)
        empty <- diverge==0
        diverge.temp <- diverge[empty]
        Z.temp <- Z[empty]
        C.temp <- C[empty]
        
        for(i in (j+1):upper){
            #do the fractal caluculation
            Z.temp <- Z.temp^2+C.temp
            diverge.temp[diverge.temp == 0 & Mod(Z.temp) >= 2] <- i
        }
        
        diverge[empty] <- diverge.temp
        Z[empty] <- Z.temp
        j <- upper
    }
    
    #perform the rendering
    diverge <- diverge+1
    maxD <- max(diverge)
    diverge.m <- matrix(diverge, n, n)
    par(mar=c(0, 0, 0, 0))
    jet.colors = colorRampPalette(c("#00007F", "blue", "#007FFF", "cyan", "#7FFF7F", 
                                    "yellow", "#FF7F00", "red", "#7F0000"))
    image(diverge.m, col=c("black", jet.colors(reps)), xlab="", ylab="")
```

---

## The good things about math

- As we see, not all math has to be boring or unattractive
- Math can be
    - Creative
    - Beautiful
    - Interactive
    - Dare I say even interesting?
- Is math hard? It can be at times, but maybe with the proper motivation others can be convinced that it's worth the effort.
- Math can even be useful every day, if someone is equipped to see the role it plays in our lives every day. Math plays a role in the form of things we see every day.
    - The arrangement of the seeds of a sunflower
    - The shape of a snowflake
    - Even the way a tree branches can be explained using math

---

## Enter the Shiny Mandelbrot App

- So to help fulfill the need to excite people about math I present the Shiny Mandelbrot App
    - http://kbelsvik.shinyapps.io/Data_Products_Project/
- This is an interactive Shiny app that can be used to create beautify images from the Mandelbrot Set
- How the math behind this is relatively simple and can be used to create some stunning images
- The math works in this way
    - We have two complex numbers Z and C
    - Z starts out equal to 0, and C begins with some value representing a point on a plot $(x, y) \rightarrow x + yi$
    - We then apply the formula $f(Z_{n+1}) = Z_{n}^2 + C$ repeatedly, counting the number of repetitions, until Z escapes into infinity (in this case we define infinity as anything greater than 2).

---
## How it all works
```{r}
z <- complex(re=0, im=0)
c <- complex(re=-0.87, im=0.25)
escape <- 0
while(Mod(z) < 2){
    z <- z^2 + c
    escape = escape+1
}
escape
```
- So now we have a number, which we map to a color, then color the the point that C represented that color, points that never escape are colored black.
- Do that for all points, and you have an image of the Mandelbrot set.
